---
title: "Factors Impacting Self-Rated Mental Health of Individuals in Canada"
subtitle: "2017 General Social Survey (GSS) for Canada"
author: "Shabnoor Nabi"
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
date: "Sept/20/2024"
output: 
  html_document:
    number_sections: yes
    toc: true
    theme: united
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```

# STUDY OVERVIEW 

Examining the effects of place of birth on individuals' self-rated mental health (while controlling for/holding constant age, gender, and marital status) based on the Canadian GSS 2017

# PACKAGES

```{r}
##Load the following packages to successfully run the code in this file

#Un-comment and install the following packages if not already loaded
# install.packages("tidyverse")
# install.packages("rio")
# install.packages("dplyr")
# install.packages("gtsummary")
# install.packages("webshot2")
# install.packages("huxtable")
# install.packages ("webshot")

#load packages
library(tidyverse)
library(rio)
library(dplyr)
library(gtsummary)
library(webshot2)
library(huxtable)
library(webshot)
```

# DATA: 2017 General Social Survey, Canada

```{r}
##Import the data file "gss-2017-CAN" 
GSS_2017 = rio::import("../Data-Raw/gss-12M0025-E-2017-c-31_F1.csv")

nrow(GSS_2017)
#The sample size without omitting the missing values across variables of interest is 20602
```

# DATA CLEANING AND MANAGEMENT

To prepare the data for analyses, first turn the values from 6-9 (where, 6 = Valid skip, 7 = Don't know, 8 = Refusal, and 9 = Not stated) within all variables of interest into NAs and rename the variables. Also, turn categorical variables into binary/dummy and classify them as factor wherever necessary.

## Outcome/Dependent Variable: Self-rated mental health (SRH_115 in codebook)


```{r}
# View the distribution of D.V.
table(GSS_2017$SRH_115) 

#Turning values 6-9 into NAs and renaming the variable as "SR_health"
GSS_2017$SR_MH <- ifelse(test = GSS_2017$SRH_115 == 6 | GSS_2017$SRH_115 == 7 | GSS_2017$SRH_115 == 8 | GSS_2017$SRH_115 == 9, yes = NA, no = GSS_2017$SRH_115)

table(GSS_2017$SR_MH)

#reverse coding the values so higher numbers indicate better self-rated health

#subsetting polcon items that need to be reverse coded
revitem = c("SR_MH")

#reverse coding the subsetted items on 1-5 scale so that the higher number represent inverse
GSS_2017[ ,revitem] = 6 - GSS_2017[ ,revitem]

##checking if reverse worked
table(GSS_2017$SR_MH)
class(GSS_2017$SR_MH)

# creating a separate factor variable for SR_health 
GSS_2017$SR_MH_R = factor(GSS_2017$SR_MH , levels = c(1,2,3,4,5) , labels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))

table(GSS_2017$SR_MH_R)
class(GSS_2017$SR_MH_R)

```


## Explanatory/Independent Variables

### A. Place of birth of respondent (BRTHCAN from code book)
```{r}
# View the distribution of the variable of interest
table(GSS_2017$BRTHCAN)

#Turning 6-9 values into NAs and renaming the variable as "PoBirth"
GSS_2017$immigrant <- ifelse(test = GSS_2017$BRTHCAN == 7 | GSS_2017$BRTHCAN == 8 | GSS_2017$BRTHCAN == 9, yes = NA, no = GSS_2017$BRTHCAN)
table(GSS_2017$immigrant)

##Recoding the above as a dummy for "Immigrant" individuals and classifying it as factor
GSS_2017$immigrant <- factor(ifelse(test = GSS_2017$immigrant == 2, yes = 1, no = 0), labels = c("Canada", "Outside Canada"))

table(GSS_2017$immigrant)
class(GSS_2017$immigrant)
```

### B. Age of the respondent at the time of survey (AGEC from code book)
```{r}
#View the distribution of the variable of interest
table(GSS_2017$AGEC)

```

### C. Sex of respondent (SEX from code book)
```{r}
#View the distribution of the variable of interest
table(GSS_2017$SEX) 

#Recoding sex as a dummy variable "SEXCAT" for 'females,' classifying it as factor, and renaming it as "female"
GSS_2017$female <- factor(ifelse(test = GSS_2017$SEX == 2, yes = 1, no = 0), labels = c("Males", "Females"))
table(GSS_2017$female)
class(GSS_2017$female)
```

### D. Marital status of the respondent (MARSTAT from code book) 
```{r}
#View the distribution 
table(GSS_2017$MARSTAT) 

#Turning 97 and 98 values into NAs and renaming the variable as "maritalstat"
GSS_2017$maritalstat <- ifelse(test = GSS_2017$MARSTAT == 97 | GSS_2017$MARSTAT == 98, yes = NA, no = GSS_2017$MARSTAT)

table(GSS_2017$maritalstat)

##Recoding the above as a dummy for "married" and common-law individuals and classifying it as factor
GSS_2017 <- GSS_2017 %>%
  mutate(marr_cohab=case_when(
    maritalstat %in% 3:6 ~ 0,
    maritalstat %in% 1:2 ~ 1))

GSS_2017$marr_cohab <- factor(GSS_2017$marr_cohab,
levels = c(0,1),
labels = c("Other", "Married/Cohabiting"))

table(GSS_2017$marr_cohab)
class(GSS_2017$marr_cohab)

```

## Sub-setting dataframe for main analyses and dropping NAs

```{r}
#Sub-setting the dataframe to variable of interest and omitting missing values across all
df_GSS2017 <- na.omit(subset(GSS_2017,  select=c(immigrant, SR_MH, SR_MH_R, female, marr_cohab, AGEC)))

nrow(df_GSS2017)
#The sample size after omitting the missing values across variables of interest is 20360
```

# DESCRIPTIVE STATISTICS

```{r}

df_GSS2017 %>% 
  select(immigrant, SR_MH, SR_MH_R, female, marr_cohab, AGEC) %>%  # keep only the columns of interest
  tbl_summary(     
    by = SR_MH,                                         # stratify entire table by outcome
    statistic = list(all_continuous() ~ "{mean} ({sd})",        # stats and format for continuous columns
                     all_categorical() ~ "{n} / {N} ({p}%)"),   # stats and format for categorical columns
    digits = all_continuous() ~ 1,                              # rounding for continuous columns
    type   = all_categorical() ~ "categorical",                 # force all categorical levels to display
    label  = list(                                              # display labels for column names
      SR_MH_R   ~ "Self-rated mental health",                           
      AGEC ~ "Age (years)",
      female    ~ "Sex",
      marr_cohab      ~ "Married/Cohabiting",
      immigrant  ~ "Place of birth"))%>%
  as_gt() |>  # convert to gt table
  gt::gtsave( # save table as image
    filename = "../Figures/descriptive_stats.png"
  )

```

# ANALYSES

Let's first do a series of simple linear bivariate analysis without any controls

A) Place of Birth & Self-rated mental health

```{r}
#Running linear regression for place of birth (I.V.) and self-rated mental health (D.V.)
mod1_BV <- lm(SR_MH ~ immigrant, data = df_GSS2017) 
# printing results
summary(mod1_BV)
```

```{r, figA, fig.height = 4, fig.width = 7, fig.align = "center"}
##Plot the association between self-rated mental health by immigration status using bar chart
FigA <- mod1_BV %>%
  ggplot(aes(SR_MH, fill = immigrant)) +
  geom_bar(position = 'dodge') +
  theme_bw() +
    labs(title = "Self-rated mental health by place of birth",
       subtitle = "GSS 2017, Canada (subset)",
       x = "Self-rated Mental Health: Poor (1) - Excellent (5)",
       y = "Frequency",
       fill = "Place of Birth")

#print barchart
FigA


#Saving fig.A as image
ggsave("../Figures/FigA.png")

##Uncomment to plot the association between self-rated mental health by immigration status using bar chart [with proportions]
# prop_table1 <- df_GSS2017 %>% 
#   group_by(immigrant, SR_MH_R) %>% 
#   tally() %>% 
#   group_by(immigrant) %>% 
#   mutate(prop = n/sum(n))
# 
# Fig1 <- ggplot(data = prop_table1, aes(immigrant, fill = SR_MH_R, y = prop)) + geom_bar(stat = "identity")  + labs(title = "Self-rated mental health by place of birth",
#                                                                                                                     fill = "Self-rated Mental Health: Excellent(5) - Poor(1)",
#                                                                                                                     subtitle = "GSS 2017, Canada (subset)",
#                                                                                                                     x = "Place of Birth",
#                                                                                                                     y = "Self-rated mental health (prop)")
# 
# #Saving fig.2 as image
# ggsave("../Figures/Fig1.png")
# 
# Fig1
```


B) Gender & Self-rated mental health

```{r}
##Running Bivariate regression for Sex (female) (I.V.) and SR_MH (D.V.)
mod2_female <- lm(SR_MH ~ female, data = df_GSS2017) 
# printing results
summary(mod2_female)
```

Let's plot and save the graph

```{r, figB, fig.height = 4, fig.width = 7, fig.align = "center"}
#Plotting self-rated mental health by SEX category using bar chart

FigB <- df_GSS2017 %>% 
  ggplot(aes(SR_MH, fill = female)) +
  geom_bar(position = 'dodge') +
  theme_bw() +
    labs(title = "Self-rated mental health by sex", 
       subtitle = "GSS 2017, Canada (subset)",
       x = "Self-rated Mental Health: Poor(1) - Excellent(5)",
       y = "Frequency",
       fill = "Sex")

#print barchart
FigB 

#Saving fig.3 as image
ggsave("../Figures/FigB.png")


##Uncomment to plot the association between self-rated mental health by sex using bar chart [with proportions]
# prop_table2 <- df_GSS2017 %>% 
#   group_by(female, SR_MH_R) %>% 
#   tally() %>% 
#   group_by(female) %>% 
#   mutate(prop = n/sum(n))
# 
# Fig2 <- ggplot(data = prop_table2, aes(female, fill = SR_MH_R, y = prop)) + geom_bar(stat = "identity")  + labs(title = "Self-rated mental health by immigrant status",
#                                                                                                                     fill = "Self-rated Mental Health: Excellent(5) - Poor(1)",
#                                                                                                                     subtitle = "GSS 2017, Canada (subset)",
#                                                                                                                     x = "Sex",
#                                                                                                                     y = "Self-rated mental health (prop)")
# 
# #Saving fig.2 as image
# ggsave("../Figures/Fig2.png")
# 
# Fig2
```



C) Marriage/Cohabitation & self rated mental health

```{r}
##Running Bivariate regression for marriage/cohabitation and self-rated mental health (D.V.)
mod3_marstat <- lm(SR_MH ~ marr_cohab, data = df_GSS2017) 
#printing results 
summary(mod3_marstat)

```

Let's plot and save the graph

```{r, echo=FALSE, figC, fig.height = 4, fig.width = 7, fig.align = "center"}

#Plotting self-rated mental health by marriage/cohabition category using bar chart

FigC <- df_GSS2017 %>% 
  ggplot(aes(SR_MH_R, fill = marr_cohab)) +
  geom_bar(position = 'dodge') +
  theme_bw() +
    labs(title = "Self-rated mental health by marital/cohab status", 
       subtitle = "GSS 2017, Canada (subset)",
       x = "Self-rated Mental Health: Poor(1) - Excellent(5)",
       y = "Frequency",
       fill = "Relationship Status")

#print barchart
FigC

#Saving fig.C as image
ggsave("../Figures/FigC.png")


##Uncomment to plot the association between self-rated mental health by marriage/cohabiting status using bar chart [with proportions]
# prop_table3 <- df_GSS2017 %>% 
#   group_by(marr_cohab, SR_MH_R) %>% 
#   tally() %>% 
#   group_by(marr_cohab) %>% 
#   mutate(prop = n/sum(n))
# 
# Fig3 <- ggplot(data = prop_table3, aes(marr-cohab, fill = SR_MH_R, y = prop)) + geom_bar(stat = "identity")  + labs(title = "Self-rated mental health by marital/Cohab status",
#                                                                                                                     fill = "Self-rated Mental Health: Excellent(5) - Poor(1)",
#                                                                                                                     subtitle = "GSS 2017, Canada (subset)",
#                                                                                                                     x = "Marital Status",
#                                                                                                                     y = "Self-rated mental health (prop)")
# 
# #Saving fig.3 as image
# ggsave("../Figures/Fig3.png")
# 
# Fig3

```

D) Age & self rated mental health

```{r, figD, fig.height = 4, fig.width = 7, fig.align = "center"}
##Running Bivariate regression for Age (I.V.) and self-rated mental health (D.V.)
mod4_age <- lm(SR_MH ~ AGEC, data = df_GSS2017) 
#printing results 
summary(mod4_age)

# Boxplot
FigD <- ggplot(data = df_GSS2017, aes(x = SR_MH_R, y = AGEC, fill = SR_MH_R)) + 
    geom_boxplot() +
    labs(title = "Self-rated mental health by age",
         x = "Self rated mental health", 
         y = "Age of respondent") +
    guides(fill = FALSE) 

#Saving fig.D as image
ggsave("../Figures/FigD.png")

FigD
```



```{r}
# Q.Does migration/immigrant status affect self-rated mental health, while holding constant marriage/cohabitation status, sex (female), and age?   

# Multivariate Regression
MVR1 <- lm(SR_MH ~ immigrant + marr_cohab + female + AGEC, data = df_GSS2017) 
#printing results of the multivariate regression
summary(MVR1)
```

```{r}
my_hux = huxreg(mod1_BV, MVR1, coefs = c(
  "(Intercept)" = "(Intercept)", 
  "Immigrant" = "immigrantOutside Canada", 
  "Married/Cohabiting" = "marr_cohabMarried/Cohabiting", 
  "Female" = "femaleFemales", 
  "Age" = "AGEC"),     
  note = "{stars}. Standard error in parentheses.", tidy_args = list(exponentiate = TRUE), bold_signif = 0.05, statistics = "BIC") %>% 
      set_caption("Bivariate (1) vs. Multivariate (2) Regression Results") %>% 
  set_align(1, everywhere, "center")

quick_html(my_hux, file = "regtable.html", open = FALSE)
# higher zoom increases resolution and image size
#run 
webshot::install_phantomjs() #for this to work
webshot::webshot(url = "regtable.html", file = "../Figures/regtable.png",
                 zoom = 5, selector = "table")

```








